Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-29 paper/main.tex

\item \texttt{require\_python(version\_spec, warn=False, extra\_msg=None, prereleases=None)}: Through \texttt{smuggle} statements and onion comments, Davos can automatically ensure that all Python packages needed to run a notebook are installed, and that the same versions of those packages are used no matter when or by whom the notebook is run.
  However, because Davos operates at runtime, one thing it cannot do automatically is install and switch to a specific version of Python itself.
  Distributing shared code along with a precise Python version for running it requires a heavier-weight solution, such as a Conda environment or Docker container (see Fig.~\ref{fig:code-sharing}).
  Yet a Davos-enhanced notebook may still \texttt{smuggle} certain packages that depend on users having a particular Python version or range of versions (e.g., even just within the standard library, the \texttt{dataclass} module was first added in Python 3.7 \cite{Smit17} and at least 19 modules are slated for removal in Python 3.13 \cite{HeimCann19}).
  The \texttt{davos.require\_python()} function can be added to the top of a Davos-enhanced notebook to communicate to users that the notebook's code should be run with a specific or constrained Python version (see Sec.~\ref{sec:illustrative-example} for example usage).
  The function may be passed a version identifier (e.g., \texttt{"3.10.5"}) or any valid version specifier \cite{CoghStuf13} (e.g., \texttt{"\raisebox{0.5ex}{\texttildelow}=3.11"}, \texttt{">=3.9;<3.12"}, etc.) and will raise an exception if the user's Python version is incompatible.
  Alternatively, a ``soft'' or suggested constraint can be imposed by passing \texttt{warn=True} to issue a warning rather than raise an error.
  Additional information can be added to the default error/warning message (e.g., the specific reason for this requirement) via the \texttt{extra\_msg} argument, and the optional \texttt{prereleases} argument can be used to explicitly allow (\texttt{True}) or disallow (\texttt{False}) pre-release versions (by default, the policy is determined by the value of \texttt{version\_spec}).

@@

After installing and importing Davos (lines 1--2), we first use the \texttt{davos.require\_python()} function to constrain the Python version used to run the notebook (see Sec.~\ref{subsec:toplevel}).
As described above, the example code in Figure \ref{fig:illustrative-example} loads two different versions of the \texttt{pandas} library: first, an older version needed to access a dataset saved in an outmoded format, then a newer one to use throughout the remainder of the notebook.
We therefore want to make sure upfront (in line 6) that the notebook's Python version falls within the range of versions that both of these two versions of \texttt{pandas} support.
%Line 6 therefore ensures upfront that the notebook's Python version falls within the overlap between the ranges of Python versions that these two versions of \texttt{pandas} support.
If it does not, the function in line 6 will raise an error that includes a message to this effect (lines 4--5).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}

@@

After smuggling the \texttt{widgetsnbextension} package in line 34, we use the \texttt{davos.use\_default\_project()} function in line 35 to revert to installing package into the notebook's project directory, restore the default value of \texttt{davos.pip\_executable} in line 36, and \texttt{smuggle} the specified version of \texttt{ipywidgets} in line 37.
With these two packages now installed
and imported, line 39 smuggles \texttt{tqdm}~\cite{daCoEtal22}, which
displays progress bars to provide status updates for running code. In
Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported
to enable more aesthetically pleasing progress bars that are displayed via
\texttt{ipywidgets}, if that package is installed and
importable. Therefore, to take advantage of this feature, it was
important to \texttt{smuggle} \texttt{tqdm} after ensuring the
\texttt{ipywidgets} package was available.

@@

Next, we load in the reformatted dataset (line 33) and pre-trained
model (line 35) that we wish to use in our analysis.  In our
hypothetical example, we can suppose that the model was provided as a
\texttt{scikit-learn} \texttt{Pipeline} object that passes data
through two pretrained models in succession.  First, a trained \texttt{CountVectorizer}
instance converts text data to an array of word counts.  Second, the
word counts are passed to a topic model~\cite{BleiEtal03} using a
pretrained \texttt{LatentDirichletAllocation} instance.
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example8}
\end{center}
Let us suppose that the \texttt{Pipeline} object had been saved by its
original creator using the \texttt{joblib} package, as
\texttt{scikit-learn}'s documentation recommends \cite{skle22}.  Because
\texttt{joblib} uses the \texttt{pickle} protocol internally, the
ability to save and load pre-trained models is not guaranteed across
different \texttt{scikit-learn} versions.  For example, suppose that
the \texttt{Pipeline} object was created using \texttt{scikit-learn}
v0.21.3.  In that version of \texttt{scikit-learn}, the
\texttt{LatentDirichletAllocation} class was defined in
\texttt{sklearn.decomposition.online\_lda}.  However, in version
0.22.0, that module was renamed to \texttt{\_online\_lda}, and in
version 0.22.1, it was again renamed to \texttt{\_lda}.

>>>>

\item \texttt{require\_python(version\_spec, warn=False, extra\_msg=None, pre\-re\-leases=\\None)}: Through \texttt{smuggle} statements and onion comments, Davos can automatically ensure that all Python packages needed to run a notebook are installed, and that the same versions of those packages are used no matter when or by whom the notebook is run.
  However, because Davos operates at runtime, one thing it cannot do automatically is install and switch to a specific version of Python itself.
  Distributing shared code along with a precise Python version for running it requires a heavier-weight solution, such as a Conda environment or Docker container (see Fig.~\ref{fig:code-sharing}).
  Yet a Davos-enhanced notebook may still \texttt{smuggle} certain packages that depend on users having a particular Python version or range of versions (e.g., even just within the standard library, the \texttt{dataclass} module was first added in Python 3.7 \cite{Smit17} and at least 19 modules are slated for removal in Python 3.13 \cite{HeimCann19}).
  The \texttt{davos.require\_python()} function can be added to the top of a Davos-enhanced notebook to communicate to users that the notebook's code should be run with a specific or constrained Python version (see Sec.~\ref{sec:illustrative-example} for example usage).
  The function may be passed a version identifier (e.g., \texttt{"3.10.5"}) or any valid version specifier \cite{CoghStuf13} (e.g., \texttt{"\raisebox{0.5ex}{\texttildelow}=3.11"}, \texttt{">=3.9;<3.12"}, etc.) and will raise an exception if the user's Python version is incompatible.
  Alternatively, a ``soft'' or suggested constraint can be imposed by passing \texttt{warn=True} to issue a warning rather than raise an error.
  Additional information can be added to the default error/warning message (e.g., the specific reason for this requirement) via the \texttt{extra\_msg} argument, and the optional \texttt{prereleases} argument can be used to explicitly allow (\texttt{True}) or disallow (\texttt{False}) pre-release versions (by default, the policy is determined by the value of \texttt{version\_spec}).

@@

After installing and importing Davos (lines 1--2), we first use the \texttt{davos.require\_\-python()} function to constrain the Python version used to run the notebook (see Sec.~\ref{subsec:toplevel}).
As described above, the example code in Figure \ref{fig:illustrative-example} loads two different versions of the \texttt{pandas} library: first, an older version needed to access a dataset saved in an outmoded format, then a newer one to use throughout the remainder of the notebook.
We therefore want to make sure upfront (in line 6) that the notebook's Python version falls within the range of versions that both of these two versions of \texttt{pandas} support.
%Line 6 therefore ensures upfront that the notebook's Python version falls within the overlap between the ranges of Python versions that these two versions of \texttt{pandas} support.
If it does not, the function in line 6 will raise an error that includes a message to this effect (lines 4--5).
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example1}
\end{center}

@@

After smuggling the \texttt{widgetsnbextension} package in line 34, we use the \texttt{davos.use\_\-default\_\-project()} function in line 35 to revert to installing package into the notebook's project directory, restore the default value of \texttt{davos.pip\_executable} in line 36, and \texttt{smuggle} the specified version of \texttt{ipywidgets} in line 37.
With these two packages now installed
and imported, line 39 smuggles \texttt{tqdm}~\cite{daCoEtal22}, which
displays progress bars to provide status updates for running code. In
Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported
to enable more aesthetically pleasing progress bars that are displayed via
\texttt{ipywidgets}, if that package is installed and
importable. Therefore, to take advantage of this feature, it was
important to \texttt{smuggle} \texttt{tqdm} after ensuring the
\texttt{ipywidgets} package was available.

@@

Next, we load in the reformatted dataset (line 33) and pre-trained
model (line 35) that we wish to use in our analysis.  In our
hypothetical example, we can suppose that the model was provided as a
\texttt{scikit-learn} \texttt{Pipeline} object that passes data
through two pretrained models in succession.  First, a trained \texttt{CountVectorizer}
instance converts text data to an array of word counts.  Second, the
word counts are passed to a topic model~\cite{BleiEtal03} using a
pretrained \texttt{LatentDirichletAllocation} instance.
\begin{center}
\includegraphics[width=0.9\textwidth]{figs/example8}
\end{center}
Let us suppose that the \texttt{Pipeline} object had been saved by its
original creator using the \texttt{joblib} package, as
\texttt{scikit-learn}'s documentation recommends \cite{skle22}.  Because
\texttt{joblib} uses the \texttt{pickle} protocol internally, the
ability to save and load pre-trained models is not guaranteed across
different \texttt{scikit-learn} versions.  For example, suppose that
the \texttt{Pipeline} object was created using \texttt{scikit-learn}
v0.21.3.  In that version of \texttt{scikit-learn}, the
\texttt{LatentDirichletAllocation} class was defined in
\texttt{sklearn.de\-comp\-o\-si\-tion.online\_lda}.  However, in version
0.22.0, that module was renamed to \texttt{\_online\_lda}, and in
version 0.22.1, it was again renamed to \texttt{\_lda}.
